\LARGE \bf Explicit Knowledge and the Deliberative Layer: Lessons Learned

### 2013-03-22 paper.tex
Added section on situation assessment

\cite{Sisbot2011}

@@

>>>>

Anchoring perceptions in a symbolic model requires perception abilities and
their symbolic interpretation. We rely on a dedicated geometric and temporal
reasoning module (as named on Fig.~\ref{fig|archi}) called SPARK (\emph{SPAtial
Reasoning \& Knowledge}~\cite{Sisbot2011}). It is a situation assessment
reasoner that generates symbolic knowledge from the geometry of the
environment with respect to relations between objects, robots and humans,
also taking into account the different perspective that each agent has on
the environment (Fig.~\ref{fig|spark-oro}).

SPARK is an \emph{amodal} geometric model of the environment that serves both
as basis for the fusion of the perception modalities and as bridge with the
symbolic layer. This geometric model is continuously updated at run-time by the
robot based on its sensors. In our experiments, objects are identified and
localised through 2D barcodes.  On the other hand, the human pose is tracked
with a Kinect-like device (assisted by motion capture to accurately track the
head motion, which is required to compute what the human is looking at).

\subsection{Computed symbolic relations}

\subsubsection{Agent-independent locations}

Absolute spatial relations (\concept{isOn}, \concept{isIn}, \concept{isNextTo},
\concept{isAbove}...) are computed for every pairs of object, taking into
account relative sizes for relations like \concept{isNextTo}. Other
agent-independent properties like \concept{isMoving} are also computed.

\subsubsection{Agent-dependent placements}

The space surrounding each agent is divided into 16 regions (like
\emph{back-left}, \emph{front-near}, etc.). Based on this partition, symbolic
relations between each pair (agent, object) is computed and added to the
knowledge base.

\subsubsection{Affordances}

SPARK also computes affordances: \concept{sees}, \concept{looksAt},
\concept{pointsAt}, \concept{canReach}. Visibility (with two levels
corresponding to the field of view and the field of attention) and pointing are
computed by placing virtual cameras in the SPARK 3D model, while reachability
relies on inverse kinematics. Several filters (based on hystersis, expected
object behaviour, etc.) allow to partially account for perception
noise~\cite{Warnier2012a}.

@@

SPARK enables \emph{perspective-taking}: spatial relations between entities can
be computed from different viewpoints, which let the robot build for each
agents it interacts with a different, \emph{perspective-aware} symbolic model
of the environment. These models are separately stored in the knowledge base.

This allows us to deal with ambiguities that arises when one speaker refers to
an object within a reference system (or changes the reference system, \ie
switches perspective) without making the reference frame
explicit~\cite{Breazeal2006, Ros2010}.

As a result, the robot stores models on the environment either in the
\emph{ego-centric} reference frame (from the robot perspective) or in
addressee-centred (\ie, \emph{allo-centric}) frames.
